---
title: 'Assignment 2: Data visualization'
author: "Tamas Nagy"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(ggrepel)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r 1. import + tidy}
tt_gh <- tt_load_gh("2020-09-22")

expeditions <- tt_download_file(tt_gh, "expeditions.csv") |> 
  select(season, peak_name)

expeditions$peak_name <- expeditions$peak_name |> 
  fct_lump_n(15)

expeditions$peak_name <- droplevels.factor(expeditions$peak_name, exclude = "Other")
  
expeditions <- na.omit(expeditions)
```

```{r 1. plotting}

expeditions |> 
  ggplot(aes(y = fct_reorder(peak_name, peak_name, .fun = length), fill = season)) +
  geom_bar() +
  scale_y_discrete(na.translate = FALSE) +
  scale_fill_viridis_d() +
  labs(title = "The 15 most popular peaks stacked by season of expedition",
       x = "Number of expeditons",
       y = NULL,
       ) +
  theme_light() +
  theme(legend.position = "bottom")

```


## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r 2. import + tidy}
tt_gh2 <- tt_load_gh("2019-02-19")

phds <- tt_download_file(tt_gh2, "phd_by_field.csv")

sum_phds <- phds |> 
  group_by(broad_field, year) |> 
  summarise(n_broad = sum(n_phds, na.rm = TRUE)) 

```

```{r 1. plotting}
sum_phds |> 
  ggplot(aes(x = year, y = n_broad, group = broad_field, color = broad_field)) +
  geom_line(size = 1.2) +
  labs(title = "Number of awarded Ph.D.-s in the US by year",
       x = NULL,
       y = NULL,
       ) +
  theme_minimal() +
  theme(legend.position = "right", aspect.ratio = 1.2) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(labels = comma_format()) +
  scale_color_brewer(palette = "Dark2", name = "Broad field")
```


## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r 3. import + tidy}
tt_gh3 <- tt_load_gh("2019-11-05")

commute <- tt_download_file(tt_gh3, "commute.csv")

commute_agg <- commute |>
  group_by(state, mode, state_abb, state_region) |> 
  summarise(n = sum(n))
  
commute_mode <- commute_agg |> 
  pivot_wider(names_from = mode, values_from = n, names_prefix = "n_") |> 
  mutate(n_Bike = coalesce(n_Bike, n_Walk),
         n_Walk = coalesce(n_Walk, n_Bike))
```

```{r 3. plotting}

commute_mode |> 
  ggplot(aes(x = n_Walk, y = n_Bike, color = state_region)) +
  geom_point(data = commute_mode[complete.cases(commute_mode$state_abb),],
             size = 2,
             shape = 16) +
  geom_text_repel(aes(label = state_abb),
                  color = "black",
                  size = 5,
                  max.overlaps = 4) +
  scale_x_log10(labels = comma_format(), limits = c(3000, 460000)) +
  scale_y_log10(labels = comma_format()) +
  labs(title = "Number of people walking vs. biking to work in each USA state",
       x = "Number of ppl walking to work (log N)",
       y = "Number of ppl biking to work (log N)",
       color = "State region"
       ) +
  theme_light() +
  theme(legend.position = "right", aspect.ratio = 1, text = element_text(size = 14))
  
```

